---
title: "Homework 3 solutions"
author: "Yuanming Mao"
output: github_document
---

```{r, include=FALSE}
library(tidyverse)
library(p8105.datasets)
library(patchwork)
library(ggridges)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Problem 1

```{r}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. 

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes. 

* How many aisles, and which are most items from?

```{r}
instacart %>% 
	count(aisle) %>% 
	arrange(desc(n))
```

There are 134 aisles and most items are from fresh vegetables aisle.

* Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.

```{r}
instacart %>% 
	count(aisle) %>% 
	filter(n > 10000) %>% 
	mutate(
		aisle = factor(aisle),
		aisle = fct_reorder(aisle, n)
	) %>% 
	ggplot(aes(x = aisle, y = n)) + 
	geom_point() + 
	theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```


* Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.

```{r}
instacart %>% 
	filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>% 
	group_by(aisle) %>% 
	count(product_name) %>% 
	mutate(rank = min_rank(desc(n))) %>% 
	filter(rank < 4) %>% 
	arrange(aisle, rank) %>% 
	knitr::kable()
```


* Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table).

```{r}
instacart %>% 
	filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
	group_by(product_name, order_dow) %>% 
	summarize(mean_hour = mean(order_hour_of_day)) %>% 
	pivot_wider(
		names_from = order_dow,
		values_from = mean_hour
	)
```


## Problem 2

* Load, tidy, and otherwise wrangle the data. Your final dataset should include all originally observed variables and values; have useful variable names; include a weekday vs weekend variable; and encode data with reasonable variable classes. Describe the resulting dataset (e.g. what variables exist, how many observations, etc).

```{r}
accel_data = 
  read.csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(weekday_vs_weekend = case_when(
    day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
    day %in% c("Saturday", "Sunday") ~ "weekend"
  )) %>% 
  relocate(week, day_id, day, weekday_vs_weekend, everything()) %>%
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    names_prefix = "activity_",
    values_to = "activity"
  ) %>% 
   mutate(
    day = factor(day),
    weekday_vs_weekend = factor(weekday_vs_weekend),
    minute = as.numeric(minute)
  ) 

``` 

The resulting dataset has `r ncol(accel_data)` variables, including week number, day ID, weekday or weekend, day name, minute of that day, and activity of that minute. There are `r accel_data %>% count(week) %>% nrow()` weeks, `r accel_data %>% count(day) %>% nrow()` days for each week, and `r accel_data %>% count(minute) %>% nrow()` minutes for each day. For each minute, we have the corresponding activity data. In total, we have `r nrow(accel_data)` observations.

* Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?

```{r}
accel_data %>%
  mutate(
    day_num = case_when(
      day == "Monday" ~ 1,
      day == "Tuesday" ~ 2,
      day == "Wednesday" ~3,
      day == "Thursday" ~4,
      day == "Friday" ~5,
      day == "Saturday" ~6,
      day == "Sunday" ~7
      ),
    day = forcats::fct_reorder(day,day_num)
  ) %>% 
  group_by(week,day) %>% 
  summarize(activity_day = sum(activity)) %>%
  pivot_wider(
    names_from = day,
    values_from = activity_day
  ) %>% 
  knitr::kable(digits = 2)

```

From the table, we can see the trend that the activity on Tuesdays and Wednesdays are quite stable, maintaining at the range from 300,000 to 470,000 counts. On other days, the activities are not that stable with some apparently lower or higher values in certain weeks.

* Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph.

```{r}
accel_data %>%
  ggplot(aes(x = minute, y = activity, color = day)) +
  geom_smooth() +
  labs(
    title = "24-hour activity time courses for each day",
    x = "Minute of day",
    y = "Activity (counts)"
  )
```

According to the graph, we can see the trend that across days of weeks, highest amounts of activities are typically around 300 to 500 counts during `r round(500/60, digit = 0)` AM to `r round(1250/60-12, digit = 0)` PM, and drop drastically until next morning. For Sunday, the activity is extremely high at around `r round(650/60, digit = 0)` AM, peaked at about 700 counts. For Friday, the activity is extremely high at around `r round(1300/60-12, digit = 0)` PM, peaked at about 800 counts.

## Problem 3

```{r}
library(p8105.datasets)
data("ny_noaa")
```

This dataset contains weather records from weather stations around the world. It contains several daily variables, including maximum and minimum temperature, total daily precipitation, snowfall, and snow depth.The dataset has `r ncol(ny_noaa)` columns and `r nrow(ny_noaa)`rows. 
`r pull(ny_noaa, prcp) %>% is.na() %>% sum()` of the stations didn't report precipitation, `r pull(ny_noaa, snow) %>% is.na() %>% sum()` of the stations didn't report snowfall,  `r pull(ny_noaa, snwd) %>% is.na() %>% sum()` of the stations didn't report snow depth, `r pull(ny_noaa, tmax) %>% is.na() %>% sum()` and `r pull(ny_noaa, tmin) %>% is.na() %>% sum()` of the stations didn't report maximum abd minumin temperature, respectively. Therefore, the resulting dataset contains extensive missing data, which could be an issue for data interpretation.

* data cleaning. Create separate variables for year, month, and day. Ensure observations for temperature, precipitation, and snowfall are given in reasonable units. 

```{r}
noaa_df = 
  ny_noaa %>% 
  separate(date, into = c("year", "month", "day"), remove = TRUE, convert = TRUE) %>% 
  mutate(
    tmin = as.numeric(tmin),
    tmax = as.numeric(tmax),
    prcp = prcp / 10,
    tmax = tmax / 10,
    tmin = tmin / 10
  )
```

* For snowfall, what are the most commonly observed values? Why?

```{r}
noaa_df %>% 
  count(snow) %>% 
  drop_na(snow) %>% 
  mutate(snow_rank = min_rank(desc(n))) %>% 
  filter(snow_rank < 4)
```

For snowfall, the most commonly observed values are 0 mm, 25 mm, and 13 mm, probably because most reported areas have no snowfall, and mild snowfalls are most common for areas that reported snowfalls.

* Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?

```{r}
noaa_df %>% 
  filter(month %in% c(1, 7)) %>% 
  mutate(month_name = case_when(
    month == 1 ~ "Janurary",
    month == 7 ~ "July"
  )) %>% 
  group_by(id, year, month_name) %>% 
  drop_na(tmax) %>% 
  summarize(mean_tmax = mean(tmax)) %>%
  ggplot(aes(x = year, y = mean_tmax, color = id)) +
  theme(legend.position = "none") +
  geom_point() +
  geom_line() +
  facet_grid(. ~ month_name) +
  labs(
    title = "Average maximum temperature in January and in July in each station across years",
    x = "Time (year)",
    y = "Mean Maxiumum temperature (C)"
  )
```

The average maximum temperatures in Janurary across years are around -10 C to 10 C, which are apparently lower than those in July that are around 20 C to 30 C. One station had an extremely low mean maximum temperature at about 14 C in July of the year around 1988.

* Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.

```{r}
tmax_tmin_plot = 
  noaa_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() + 
  labs(
    title = "Daily minimum temperature vs minimum temperature",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)"
  )

snow_plot = 
  noaa_df %>% 
  filter(snow > 0 & snow < 100) %>% 
  mutate(year = as.character(year)) %>% 
  ggplot(aes(x = snow, y = year)) +
  geom_density_ridges() +
  labs(
    title = "Distribution of snowfall values greater than 0 and less than 100 by year",
    y = "Time (year)",
    x = "Snowfall (mm)"
    )

tmax_tmin_plot + snow_plot

```

